---
title: "STAT 421 \n Assignment #7"
author: "Duncan Gates"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  prettydoc::html_pretty:
    theme: tactile # Other options are cayman, tactile, architect, leonids, hpstr
    highlight: vignette
    math: katex #Instead of mathjax so it works offline
documentclass: book
classoption:
  - twocolumn
  - landscape
papersize: a5
linestretch: 1.5
fontsize: 12pt
links-as-notes: true
header-includes:
  - \usepackage{indentfirst}
  - \usepackage{amsmath}
---

```{r setup, include=FALSE}
# library(pagedown)
# setwd("/Users/dunk/Classes/STAT421/Assignment 7")
# pagedown::chrome_print(input = "Assignment-7.html",
#                       output = "Assignment7.pdf")
```


## 4.56

In 4.7 the 30 minute arrival time has a unform distribution therefore we must find, 

$$
\begin{aligned}
P(25 < Y < 30|Y > 10)&=\frac{\frac{1}{6}}{\frac{2}{3}} \\
&= \boxed{\frac{1}{4}}
\end{aligned}
$$

## 4.59 

I just typed all of these into my TI-83 hope that's alright.

**a.**

$$
\begin{aligned}
P(Z >z_0)&=0.5 \\
z_0&=\boxed{0.5}
\end{aligned}
$$

**b.**

$$
\begin{aligned}
P(Z <z_0)&=0.8643 \\
z_0&=\boxed{1.1}
\end{aligned}
$$

**c.**

$$
\begin{aligned}
P(-z_0<Z <z_0)&=0.9 \\
z_0&=\boxed{1.645}
\end{aligned}
$$

**d.**

$$
\begin{aligned}
P(-z_0<Z <z_0)&=0.99 \\
z_0&=\boxed{2.576}
\end{aligned}
$$


## 4.74

**a.**

$$
\begin{aligned}
P(Y >72)&=P(\frac{y-\mu}{\sigma}) \\
&=P(\frac{72-78}{6}) \\
&=P(z \ge -1) \\
&=\boxed{0.8413}
\end{aligned}
$$

**b.**

$$
\begin{aligned}
0.9&=P(z \ge \frac{x-78}{6}) \\
1.28&=\frac{x-78}{6} \\
x&=\boxed{85.68}
\end{aligned}
$$

**c.**

$$
\begin{aligned}
0.281&=P(z \ge \frac{x-78}{6}) \\
0.58&=\frac{x-78}{6} \\
x&=\boxed{81.48}
\end{aligned}
$$

**d.**

$$
\begin{aligned}
0.25&=P(z \ge \frac{x-78}{6}) \\
0.67&=\frac{x-78}{6} \\
x&=73.98 \\
P(Y>78.98)&=P(z \ge \frac{78.98-78}{6}) \\
&=\boxed{0.4369}
\end{aligned}
$$

**f.**

$$
\begin{aligned}
P(Y > 84|Y >72)&=\frac{P(Y>84)}{P(Y>72)} \\
&=\frac{P(z > 1)}{P(z > -1)} \\
&= \frac{0.1587}{0.8413} \\
&=\boxed{0.1886}
\end{aligned}
$$


## 4.88

**a.**

$$
\begin{aligned}
P(Y > 3)&=1-P(Y \le 3) \\
&=1-\int_0^3f(y)\:dy \\
&=1-\int_0^3(\frac{1}{2.4})e^{\frac{-y}{2.4}}\:dy \\
&=1-\frac{2.4}{2.4}\Bigg[-e^{-3/2.4}-(-e^{-0/2.4})\Bigg] \\
&=\boxed{e^{-1.25}}
\end{aligned}
$$

**b.**

$$
\begin{aligned}
P(2 < y < 3)&= \int_2^3(\frac{1}{2.4})e^{\frac{-y}{2.4}}\:dy \\
&=\frac{2.4}{2.4}\Bigg[-e^{-3/2.4}-(-e^{-2/2.4})\Bigg] \\
&=e^{-1.25}-(-e^{-.83}) \\
&=\boxed{\approx0.1481}
\end{aligned}
$$

## 4.92

Given $C=100+40Y+3Y^2$

We find $E(Y)$ as 

$$
\begin{aligned}
E(Y)&=\frac{1}{\mu} \\
&=\frac{1}{(1/10)} \\
&=10
\end{aligned}
$$
and $E(Y^2)$

$$
\begin{aligned}
E(Y^2)&=\int_0^{\infty}\frac{y^2}{10}e^\frac{-y}{10}dy \\
&=\frac{1}{10}\Bigg(\frac{\tau(3)}{(1/10)^3}\Bigg) \\
&=200 \\
E(Y^3)&=\frac{1}{10}\Bigg(\frac{\tau(4)}{(1/10)^4}\Bigg) \\
&=6000 \\
E(Y^4)&=\frac{1}{10}\Bigg(\frac{\tau(3)}{(1/10)^5}\Bigg) \\
&=240000
\end{aligned}
$$

$$
\begin{aligned}
E(C)&=E(100+40Y+3Y^2) \\
&=E(100)+40*E(Y)+3*E(Y^2) \\
&=100+40*10+3*200 \\
&=\boxed{1100}
\end{aligned}
$$

$$
\begin{aligned}
V(C)&=E(C^2)-E(C)^2 \\
&=E(100+40Y+3Y^2)^2-(1100)^2 \\
&=E((100)^2+(40Y)^2+(3Y^2)^2+2(3Y^2*40Y)+2(40Y*100)+2(3Y^2*100))-(1100)^2 \\
&=4,130,000-(1,100)^2 \\
&=\boxed{2,920,000}
\end{aligned}
$$

## 4.96

**a.**

We have $\alpha = 4$ and $\beta = 2$ therefore,

$$
\begin{aligned}
1 &= k\int_0^{\infty}y^3e^{-y/2}\:dy\\
k&= \frac{1}{\tau(4)2^4} \\
&=\boxed{\frac{1}{96}}
\end{aligned}
$$

**b.**

$$
\begin{aligned}
f(y)&=\frac{y^3e^{-y/2}}{96} \\
&=\frac{y^{4-1}e^{-y/2}}{(2^4)(4-1)!} \\
&=\frac{y^{\frac{8}{2}-1}e^{-y/2}}{(2^\frac{8}{2})\tau(\frac{8}{2})} \\
&\text{and} f(\chi^2)=\frac{y^{\frac{n}{2}-1}e^{-y/2}}{(2^\frac{n}{2})\tau(\frac{n}{2})}
\end{aligned}
$$

$$
\begin{aligned}
n&=2 \beta \\
n&=8
\end{aligned}
$$

So yes with 8 degrees of freedom.

**c.**

The mean can be calculated as

$$
\begin{aligned}
E(Y)&=\alpha \beta \\
&=\boxed{8}
\end{aligned}
$$

The variance is 

$$
\begin{aligned}
V(Y)&=\alpha \beta^2 \\
&=\boxed{16}
\end{aligned}
$$

**d.**

$$
\begin{aligned}
\sigma&=\sqrt{V(X)} \\
&=\sqrt{16} \\
P(|Y-8|<2(4))&=P(0<Y<16) \\
&=0.95762
\end{aligned}
$$

## 4.126

**a.**

$$
\begin{aligned}
F(y)&=P(Y \le y) \\
&=\int_0^y(6t-6t^2)\:dt \\
&=3y^2-2y^3
\end{aligned}
$$

So we have the cumulative probability function 

$$
\begin{aligned}
F(y)&= 
\left\{
\begin{array}{ll}
      0 & y < 0 \\
      3y^2-2y^3 & 0\leq y\leq 1 \\
      1 & y > 1 \\
\end{array} 
\right.
\end{aligned}
$$

**b.**

```{r, echo = F}
library(ggplot2)
f <- function(x){ 
  return(ifelse(0<= x & x<=1, 3*x^2-2*x^3,
                        ifelse(x> 1, 1, 
                               ifelse(x< 0, 0, 0))))}
ggplot2::ggplot(data.frame(x = c(-1.25,1.25)), aes(x = x)) + stat_function(fun = f) + ggtitle("F(y)") + labs(x = "y", y = "F(y)")
```

```{r, echo = F}
p <- function(x){ return(6*x-6*x^2)}
ggplot2::ggplot(data.frame(x = c(0,1)), aes(x = x)) + stat_function(fun = p) + ggtitle("f(y)") + labs(x = "y", y = "f(y)")
```


**c.**

$$
\begin{aligned}
P(0.5 \le Y \le 0.8)&=F(0.8)-F(0.5) \\
&=1.92-1.092-0.75+0.25 \\
&= \boxed{0.396}
\end{aligned}
$$

## 4.134


**a.**

Given $\alpha = 4$ and $\beta = 7$,

$$
\begin{aligned}
P(Y \le 0.7)&=F(0.7) \\
&=\sum_{i = 4}^{10}{10 \choose i}(0.7)^i(0.3)^{10-i} \\
&=P(4 \le X \le 10) && \text{distributed binomially n = 10 and p = 0.7} \\
&=\boxed{0.989}
\end{aligned}
$$

**b.**

$$
\begin{aligned}
P(Y \le 0.6)&=F(0.6) \\
&=\sum_{i = 4}^{10}{10 \choose i}(0.6)^i(0.4)^{10-i} \\
&=P(12 \le X \le 25) && \text{distributed binomially n = 25 and p = 0.6} \\
&=\boxed{0.922}
\end{aligned}
$$

## 4.142

**a.**

$$
\begin{aligned}
m_Y(t)&=E(e^{ty}) \\
&=\int_0^1e^{ty}\:dy \\
&=\boxed{\frac{e^t-1}{t}}
\end{aligned}
$$

**b.**

Given $W=aY$

$$
\begin{aligned}
m_W(t)&=E(e^{tw}) \\
&=m_Y(at) \\
&=\boxed{\frac{e^{at}-1}{at}}
\end{aligned}
$$

**c.**

Given $W=-aY$,

$$
\begin{aligned}
m_W(t)&=E(e^{tw}) \\
&=m_Y(-at) \\
&=\boxed{\frac{1-e^{-at}}{at}}
\end{aligned}
$$

**d.**

Given $V=aY+b$

$$
\begin{aligned}
m_V(t)&=E(e^{tv}) \\
&=E(e^{t(ay+b)}) \\
&=e^{bt}E(e^{aty}) \\
&=e^{bt}*m_Y(at) \\
&=\boxed{\frac{e^{bt}-e^{(a+b)t}}{at}}
\end{aligned}
$$

## 4.190

**a.**

$$
\begin{aligned}
r(t)&=\frac{f(t)}{1-F(t)} \\
&=\frac{\lambda e^{-\lambda t}}{1-1+e^{-\lambda t}} \\
&=\lambda
\end{aligned}
$$


**b.**

For a Weibull function with $m > 1$,

$$
\begin{aligned}
r(t)&=\frac{\frac{my^{m-1}}{\alpha}e^{-y^m/ \alpha}}{1-1+e^{-y^m/ \alpha}} \\
&= \frac{mt^{m-1}}{\alpha}
\end{aligned}
$$

Therefore $r(t)$ is an increasing function of t when $m > 1$























